Nanophotonic Inverse Design

### 2011-12-14 presentations/inv_dev_plan_dec-2011/paper.tex
Draft of theory notes for admm finished.

\begin{appendix}
\section{Constructing the relevant matrices and vectors}
The physics residual is based on the electromagnetic wave equation, 
    \begin{equation}
    \nabla \times \epsilon^{-1} \nabla \times H - \mu \omega^2 H - 
        \nabla \times \epsilon^{-1} J - i \omega M = 0,
    \end{equation}
    which can be written as
    \begin{equation}
    r(x,p) = A(p) x - b(x) = B(x) p - d(x).
    \end{equation}

\subsection{Defining the curl operators}
Specifically, the curl operations are defined as
    \begin{equation}
    \nabla^\pm \times u = \begin{bmatrix}
        D_y^\pm u_z - D_z^\pm u_y \\
        D_z^\pm u_x - D_x^\pm u_z \\
        D_x^\pm u_y - D_y^\pm u_x
        \end{bmatrix}
    \end{equation}
    where
    \begin{subequations}\begin{align}
    D_x^- v &= \frac{v(i,j,k) - v(i-1,j,k)}{\phi_x(i-\frac{1}{2})} \\
    D_x^+ v &= \frac{v(i+1,j,k) - v(i,j,k)}{\phi_x(i+\frac{1}{2})},
    \end{align}\end{subequations}
    and likewise for the numerical derivatives in the $y$ and $z$ directions.
Note that we assume that $\Delta x, \Delta y, \Delta z = 1$, and so omit them
    from the denominators of $D$.

Lastly, the $\phi$ functions are used to construct 
    absorbing boundary conditions.
Specifically, we use stretched-coordinate perfectly matched layers\cite{scpml}
    for which $\phi$ is defined as
    \begin{equation}
    \phi_x(x) = \begin{cases}
        1 + \frac{i\sigma}{\omega} x^m & x < t_\text{PML}, \\ 
        1 + \frac{i\sigma}{\omega} (x_\text{max} -x)^m
            & x > x_\text{max} - t_\text{PML}, \\ 
        1 & \text{otherwise,}
        \end{cases}
    \end{equation}
    where $\sigma$ is the strength of the absorbing layer, 
    and $m$ controls the profile of the layer.
Typical values for $\sigma$ and $m$ are $1$ and $2.5$ respectively.

\subsection{Forming the field subproblem}
$A(p)$ and $b(p)$ are constructed as follows,
    \begin{subequations}\begin{align}
    A(p) &= A_1 \diag(\epsilon(p)^{-1}) A_2 - \mu \omega^2 I \\
    b(p) &= A_1 \diag(\epsilon(p)^{-1}) b_j + i \omega b_m \\
    A_1 &= \nabla^- \times \\
    A_2 &= \nabla^+ \times \\
    b_j &= J \\
    b_m &= M,
    \end{align}\end{subequations}
    where $\nabla^+$ and $\nabla^-$ refer to forward- and backward-differencing
    respectively.

\subsection{Forming the structure subproblem}
We now show how to form $r(x,p) = B(x)p - d(x)$.

\section{Solving the matrices}
\section{Linear algebra definitions}
\subsection{Gradient and Hessian calculations}
Mention something about complex values.
\subsection{Matrix inversion lemma}\label{inv lemma}
\begin{equation}
(A + UCV)^{-1} = A^{-1} - A^{-1}U (C^{-1} + V A^{-1} U)^{-1} V A^{-1}
\end{equation}
\subsection{Block elimination of a matrix}\label{block elim}
\begin{equation}
\begin{bmatrix} A_{11} & A_{12} \\ A_{21} & A_{22} \end{bmatrix} 
\begin{bmatrix} x_1 \\ x_2 \end{bmatrix} = 
\begin{bmatrix} b_1 \\ b_2 \end{bmatrix}
\end{equation}
\begin{subequations}\begin{align}
A_{11} x_1 + A_{12} x_2 &= b_1 \\
x_1 &= A_{11}^{-1} (b_1 - A_{12} x_2)
\end{align}\end{subequations}
\begin{subequations}\begin{align}
A_{21} x_1 + A_{22} x_2 &= b_2 \\
A_{21} A_{11}^{-1} (b_1 - A_{12} x_2) + A_{22} x_2 &= b_2 \\
(A_{22} - A_{21} A_{11}^{-1} A_{12}) x_2 &= b_2 - A_{21} A_{11}^{-1} b_1
\end{align}\end{subequations}
Typically, one computes $A_{11}^{-1} A_{12}$ and $A_{11}^{-1} b_1$, 
    and then solves for the rest of the system.
\end{appendix}
\begin{thebibliography}{99}
\bibitem{ADMM} Boyd group ADMM paper
\bibitem{scpml} \url{http://math.mit.edu/~stevenj/18.369/pml.pdf}
\end{thebibliography}
\end{document}

>>>>

\begin{appendix}
\section{Constructing the relevant matrices and vectors}
The physics residual is based on the electromagnetic wave equation, 
    \begin{equation}
    \nabla \times \epsilon^{-1} \nabla \times H - \mu \omega^2 H - 
        \nabla \times \epsilon^{-1} J - i \omega M = 0, \label{em wave}
    \end{equation}
    which can be written as
    \begin{equation}
    r(x,p) = A(p) x - b(x) = B(x) p - d(x).
    \end{equation}

\subsection{Defining the curl operators}
The various curl operators in \eqref{em wave} do not all
    correspond to the same matrix.
We use both forward-differencing $(\nabla^+)$
    and backward-differencing $(\nabla^-)$ curl operators.
These are defined as
    \begin{subequations}\begin{align}
    \nabla^\pm \times u &= \begin{bmatrix}
        0 & -S_z^\pm D_z^\pm & S_y^\pm D_y^\pm \\
        S_z^\pm D_z^\pm & 0 & -S_x^\pm D_x^\pm \\
        -S_y^\pm D_y^\pm & S_x^\pm D_x^\pm & 0 \end{bmatrix}
        \begin{bmatrix} u_x \\ u_y \\ u_z \end{bmatrix}
    \\
    &= \begin{bmatrix}
        S_y^\pm D_y^\pm u_z - S_z^\pm D_z^\pm u_y \\
        S_z^\pm D_z^\pm u_x - S_x^\pm D_x^\pm u_z \\
        S_x^\pm D_x^\pm u_y - S_y^\pm D_y^\pm u_x
        \end{bmatrix}
    \end{align}\end{subequations}
    where the $D\pm$ matrices perform either forward- or backward-differencing,
    and the $S^\pm$ matrices implement the stretched-coordinate 
    absorbing boundary conditions\cite{scpml}.

These matrices are defined as follows,
    \begin{subequations}\begin{align}
    D_x^+ u_w &= \frac{u_w(i+1,j,k) - u_w(i,j,k)}{\Delta x} \\
    D_x^- u_w &= \frac{u_w(i,j,k) - u_w(i-1,j,k)}{\Delta x} \\
    S_x^+ v &= \frac{v(i,j,k)}{\phi_x(i+\frac{1}{2})} \\
    S_x^- v &= \frac{v(i,j,k)}{\phi_x(i)}, 
    \end{align}\end{subequations}
    and likewise for the $y$ and $z$ directions.
Here, the $w$ subscript signifies either $x$, $y$, or $z$,
    while $v$ is just a placeholder vector.
Also, the $\phi$ functions are defined according to \cite{scpml} as
    \begin{equation}
    \phi_x(x) = \begin{cases}
        1 + \frac{i\sigma}{\omega} x^m & x < t_\text{PML}, \\ 
        1 + \frac{i\sigma}{\omega} (x_\text{max} -x)^m
            & x > x_\text{max} - t_\text{PML}, \\ 
        1 & \text{otherwise,}
        \end{cases}
    \end{equation}
    where $\sigma$ is the strength of the perfectly-matched layer (PML), 
    $m$ controls the profile of the PML,
    and $t_\text{PML}$ is the thickness of the PML.
Typical values for $\sigma$, $m$, and $t_\text{PML}$ are 
    $1.0$, $2.5$, and $10$ respectively.

Lastly, the $D$ matrices impose periodic (wrap-around) boundary conditions
    for cells adjacent to the boundary.
Also, note that the assymetry in the definition of the $S^\pm$ matrices
    is due to the structure of the Yee cell\cite{yee cell}.

\subsection{Forming the field subproblem}
$A(p)$ and $b(p)$ are constructed as follows,
    \begin{subequations}\begin{align}
    A(p) &= A_1 \diag(\epsilon(p)^{-1}) A_2 - \mu \omega^2 I \\
    b(p) &= A_1 \diag(\epsilon(p)^{-1}) b_j + i \omega b_m, \\
    \end{align}\end{subequations}
    where
    \begin{subequations}\begin{align}
    A_1 &= \nabla^+ \times \\
    A_2 &= \nabla^- \times \\
    b_j &= J \\
    b_m &= M.
    \end{align}\end{subequations}

\subsection{Forming the structure subproblem}
We now show how to form $r(x,p) = B(x)p - d(x)$.
    \begin{subequations}\begin{align}
    r(x,p) &= A(p)x - b(p) \\
        &= (A_1 \diag(\epsilon(p)^{-1}) A_2 - \mu \omega^2 I) x - 
            (A_1 \diag(\epsilon(p)^{-1}) b_j + i \omega b_m) \\ 
        &= A_1 \diag(\epsilon(p)^{-1}) (A_2 x - b_j) - 
            (\mu \omega^2 x + i \omega b_m) \\ 
        &= A_1 \diag(A_2 x - b_j) \epsilon(p)^{-1} - 
            (\mu \omega^2 x + i \omega b_m),
    \end{align}\end{subequations}
    and choosing $\epsilon(p)^{-1} = A_3 p + p_0$, we obtain
    \begin{subequations}\begin{align}
    r(x,p) &= A_1 \diag(A_2 x - b_j) (A_3 p + p_0) - 
            (\mu \omega^2 x + i \omega b_m) \\
        &= B(x) p - d(x),
    \end{align}\end{subequations}
    where
    \begin{subequations}\begin{align}
    B(x) &= A_1 \diag(A_2 x - b_j) A_3\\
    d(x) &= \mu \omega^2 x + i \omega b_m - A_1 \diag(A_2 x - b_j) p_0.
    \end{align}\end{subequations}

\section{Solving the matrices}
\section{Linear algebra definitions}
\subsection{Gradient and Hessian calculations}
The following convention is assumed throughout these notes.
Given $f(x) \in \reals^n \to \reals$,
    \begin{equation}
    \nabla f(x) = \begin{bmatrix}
        \frac{df}{dx_1} \\
        \frac{df}{dx_2} \\
        \vdots
        \end{bmatrix}
    \end{equation}
    and
    \begin{equation}
    \nabla^2 f(x) = \begin{bmatrix}
        \frac{d^2f}{dx_1^2} & \frac{d^2f}{dx_1 dx_2} & \cdots \\
        \frac{d^2f}{dx_2 dx_1} & \frac{d^2f}{dx_2^2} & \cdots \\
        \vdots & \vdots & \ddots
        \end{bmatrix}.
    \end{equation}

\subsection{Handling complex values}

\subsection{Matrix inversion lemma}\label{inv lemma}
The matrix inversion simply states that
\begin{equation}
(A + UCV)^{-1} = A^{-1} - A^{-1}U (C^{-1} + V A^{-1} U)^{-1} V A^{-1}.
\end{equation}

\subsection{Block-elimination of a matrix}\label{block elim}
Solving $Ax=b$ via block-elimination involves first splitting the problem up as
    \begin{equation}
    \begin{bmatrix} A_{11} & A_{12} \\ A_{21} & A_{22} \end{bmatrix} 
    \begin{bmatrix} x_1 \\ x_2 \end{bmatrix} = 
    \begin{bmatrix} b_1 \\ b_2 \end{bmatrix},
    \end{equation}
    solving for $x_1$ in terms of $x_2$, 
    \begin{subequations}\begin{align}
    A_{11} x_1 + A_{12} x_2 &= b_1 \\
    x_1 &= A_{11}^{-1} (b_1 - A_{12} x_2), \label{x1 solve}
    \end{align}\end{subequations}
    substituting in order to solve for $x_2$,
    \begin{subequations}\begin{align}
    A_{21} x_1 + A_{22} x_2 &= b_2 \\
    A_{21} A_{11}^{-1} (b_1 - A_{12} x_2) + A_{22} x_2 &= b_2 \\
    (A_{22} - A_{21} A_{11}^{-1} A_{12}) x_2 &= b_2 - A_{21} A_{11}^{-1} b_1
    \end{align}\end{subequations}
    and then plugging the solution in to \eqref{x1 solve} to back out $x_1$.
Typically, one pre-computes $A_{11}^{-1} A_{12}$ and $A_{11}^{-1} b_1$, 
    and then solves for the rest of the system.

Additionally, if we are only concerned about the value of $x_1$,
    we can simply use the formula
    \begin{equation}
    x_1 = A_{11}^{-1} (b_1 - A_{12} 
        (A_{22} - A_{21} A_{11}^{-1} A_{12})^{-1}
        (b_2 - A_{21} A_{11}^{-1} b_1)).
    \end{equation}
\end{appendix}
\begin{thebibliography}{99}
\bibitem{ADMM} Boyd group ADMM paper
\bibitem{scpml} \url{http://math.mit.edu/~stevenj/18.369/pml.pdf}
\end{thebibliography}
\end{document}
